% resume.tex
%
% (c) 2011 Alex Kunze Susemihl <alexsusemihl@gmail.com>
% $Id: resume.tex,v 1.1 28/06/2008 13:36:06 $
%
\documentclass[10pt]{article}
\usepackage{fullpage}
\usepackage[scaled]{helvet}
%\usepackage{microtype}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[latin1]{inputenc}
\usepackage[brazil]{babel}
\textheight=10.0in
\pagestyle{plain}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}
\setlength{\textwidth}{6.5in}
\begin{document}
\thispagestyle{empty}
\begin{tabular*}{6.5in}{l@{\extracolsep{\fill}}r}
\LARGE{\bf Alex Kunze Susemihl}& \\
&\\
\footnotesize{Ahlbecker Stra{\ss}e, 6}&\footnotesize{+49-173-4682162} \\
\footnotesize{10437 Berlin, Germany}&\footnotesize{+49-30-68321681}\\
&\footnotesize{alexsusemihl@gmail.com} \\
\end{tabular*}
\\
\vspace{0.2in}

\vspace{0.15in}
{\large \textbf{Pre-Publication and Contribution}}
\vspace{0.3in}

\hspace{0.1in}\parbox[l]{6.4in}{	{\bf Susemihl, A.}, Meir, R., and Opper, M. (2011). \emph{Analytical Results for the Error in Filtering of Gaussian Processes}. In Shawe-Taylor, J., Zemel, R. S., Bartlett, P., Pereira, F., and Weinberger, K., editors, Advances in Neural Information Processing Systems 24, pages 2303 - 2311.}
\vspace{0.1in}

The research project was conceived by Manfred Opper and Ron Meir with my contribution. The research was carried out by me under Manfred Opper's supervision an counselling. The
paper was written by me with input from Manfred Opper and Ron Meir under their supervision and advice.\par

\vspace{0.15in}

\hspace{0.1in}\parbox[l]{6.4in}{	{\bf Susemihl, A.}, Meir, R., and Opper, M. (2013) \emph{Dynamic State Estimation Based on Poisson Spike Trains: Towards an Optimal Theory of Encoding.} Journal of Statistical Mechanics: Theory and Experiment, http://dx.doi.org/10.1088/1742-5468/2013/03/P03009}
\vspace{0.1in}

The research project was conceived by me, Manfred Opper and Ron Meir. The research was carried out by me under Manfred Opper's and Ron Meir's supervision. The paper was written
mainly by me, with revision and supervision from Manfred Opper and Ron Meir.

\vspace{0.15in}
\hspace{0.1in}\parbox[l]{6.4in}{{\bf Susemihl, A.}, Meir, R., Opper, M. (2014), \emph{Optimal Neural Codes for Control and Estimation}. Submitted to {\em Neural Information Processing Systems 2014}, available from \emph{http://arxiv.org/pdf/1406.7179v1.pdf}}
\vspace{0.1in}

The research project was conceived by Ron Meir, Manfred Opper and me. The research was carried out by me under Ron Meir's and Manfred Opper's supervision. The paper was writen
by me and Ron Meir with Manfred Opper's supervision and revision.

\vspace{3.5in}

\begin{flushright}
\Large{Alex Kunze Susemihl}
\end{flushright}

\end{document}
